Age-related changes in plant growth and traits

### 2015-05-21 results_accessory_tissues.Rmd
lots of standardizing names also recalculated accessory costs on a per seed basis and added figures

###How do accessory costs shift with seed size?
- Larger-seeded species have higher accessory costs. 
- Also shown in other studies and suggested as a mechanism for why angiosperms were able to have such small seeds (relative to gymnosperms, with no decrease in proportional accessory costs - with decreasing size)  
```{r, echo=FALSE,message=FALSE}
yvar <- "PerAcc_mean"
xvar <- "seed_size"
data <- SummarySpp

plot_yvar_vs_xvar(data, yvar, xvar,log="x",
                  xlab=xvar, ylab=yvar)
mod <- lm(log(seed_size)~PerAcc_mean,SummarySpp)
summary(mod)
anova(mod)
```

###Do species differ in the proportional energy allocated to propagules (vs. accessory tissues)  
- COER, EPMI, HEPU, LEES, PILI invest the most energy in propagules  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPropagule"
xvar <- "species"
data <- filter(SummaryInd,Total>0)

plot_yvar_vs_xvar(data, yvar, xvar,
                 xlab=xvar, ylab=yvar)
mod <- lm(asin(sqrt(PerPropagule)) ~ species-1, data)
summary(mod)
data.frame(
    select(glance(mod), r.squared, p.value),
    n= length(resid(mod)),
    a = coef(mod)[1],
    b = coef(mod)[2]
  )

tmp <- summary(mod)
tmp <- tmp$coefficients
tmp <-as.data.frame(tmp)
tmp <- cbind(names(labels.spp),tmp)
names(tmp) <- c("species","estimate.trans","se.trans","t.value","p.value")
tmp2 <- select(SummarySpp, species, PerPropagule_mean, PerPropagule_se)
tmp <- merge(tmp, tmp2, by="species")
tmp
plot(PerPropagule_mean~species,tmp,cex.axis=.8,pch=16,log="y",ylim=c(.001,.25),main="Percent reproductive investment put into propagules")

mean1 <- tmp$PerPropagule_mean
se1 <- tmp$PerPropagule_se

@@

#for(i in 1:14) {
#arrows(i,tmp$PerPropagule_mean[i]+tmp$PerPropagule_se[i],i,tmp$PerPropagule_mean[i]-tmp$PerPropagule_se[i],code=3,angle=90, length=0.1)}

#barplot(height=test$PerPropagule_mean, names=test$species,cex.axis=.8,ylim=c(0,.25),cex.names=.8)
#xv <- barplot(height=test$PerPropagule_mean, names=test$species,cex.axis=.8,ylim=c(0,.25),cex.names=.8)
#xv

#points((PerPropagule_mean+PerPropagule_se)~species,test,pch=25)
#points((PerPropagule_mean-PerPropagule_se)~species,test,pch=24)

@@

###Do species differ in the proportional energy allocated to aborted pre-pollination tissues
- BOLE, EPMI, GRSP, HATE invest the most energy in aborted pre-pollination tissues  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "species"
data <- filter(SummaryInd,Total>0 & age>2)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar)
mod <- lm(asin(sqrt(PerPrePol_A)) ~ species, data)
summary(mod)
```

##Do individual species decrease accessory costs as they increase in size, age or RA?
###Change in accesory costs With size:  
- PerAcc is arcsine transformed, size is log transformed  
- COER, HATE, and PILI show a significant decrease in accessory costs with increasing size (p<0.05)  
- GRSP shows a marginally significant decrease in accessory costs with increasing size (0.05<p<0.10)  
- PEPU shows a marginally significant INCREASE in accessory costs with increasing size (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "FinalWeight"

@@

###Change in accesory costs With age:
- PerAcc is arcsine transformed, age is log transformed  
- BOLE, COER, GRSP, HATE, HEPU, PELA show a significant decrease in accessory costs with increasing age (p<0.05)  
- EPMI, GRBU shows a marginally significant decrease in accessory costs with increasing age (0.05<p<0.10)  
- PUTU shows a sig INCREASE in accessory costs with increasing age  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerAcc"
xvar <- "age"

@@

###Change in accesory costs With respect to RA:  
- both RA and PerAcc are arcsine transformed  
- BAER, COER, GRBU, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in accessory costs with increasing RA (p<0.05)  
- BOLE shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPropagule"
xvar <- "RA"

@@

##Next considering if the amount of energy spent on aborted buds shifts with age, size, RA  
###Change in aborted bud investment with size:  
- PerPrePol_A is arcsine transformed, size is log transformed  
- GRSP, HATE, HEPU show a significant decrease in aborted buds with increasing size (p<0.05)  
- PUTU shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "FinalWeight"

@@

###Change in aborted bud investment with respect to age:  
- PerPrePol_A is arcsine transformed, age is log transformed  
- BAER, BOLE, COER, EPMI, LEES, PHPH, PILI show a significant decrease in aborted buds with increasing age (p<0.05)  
- GRSP, HATE show a marginally significant decrease in aborted buds with increasing age (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "age"

@@

###Change in aborted bud investment with respect to RA:
- both RA and PerPrePol_A are arcsine transformed  
- BAER, BOLE, COER, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in aborted buds with increasing RA (p<0.05)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_A"
xvar <- "RA"

@@

##Do individual species decrease total pre-pollination costs as they increase in size, age or RA?  
###Do total pre-pollination costs change with respect to size:  
- PerPrePol_All is arcsine transformed, size is log transformed  
- GRSP, HATE show a significant decrease in total pre-pollination costs with increasing size (p<0.05)  
- HEPU shows a marginally significant decrease in total pre-pollination costs with increasing size (0.05<p<0.10)

```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "FinalWeight"

@@

###Do total pre-pollination costs change with respect to age:
- PerPrePol_All is arcsine transformed, age is log transformed  
- BAER, BOLE, COER, HEPU, LEES, PILI show a significant decrease in total pre-pollination costs with increasing age (p<0.05)  
- EPMI, GRBU, GRSP, HATE shows a marginally significant decrease in total pre-pollination costs with increasing age (0.05<p<0.10)  
- PUTU shows a sig INCREASE in total pre-pollination  costs with increasing age  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "age"

@@

### Do total pre-pollination costs change with respect to RA:  
- both RA and PerPrePol_All are arcsine transformed    
- BAER, BOLE, COER, GRSP, HATE, HEPU, LEES, PEPU show a significant decrease in total pre-pollination costs with increasing RA (p<0.05)  
- PELA shows a marginally significant decrease in total pre-pollination costs with increasing RA (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "RA"

@@

### Do total pre-pollination costs change with respect to Reproductive Investment:
- PerPrePol_All is arcsine transformed, ReproInv is log transformed  
- COER, GRBU, GRSP, HATE, PEPU, PHPH, PILI show a significant decrease in total pre-pollination costs with increasing reproductive investment (p<0.05)  
- LEES shows a marginally significant decrease in total pre-pollination costs with increasing reproductive investment (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "PerPrePol_All"
xvar <- "ReproInv"

@@

###For each species, look at pattern of investment in different tissues across ages  
- Most obvious pattern is that different species invest more or less in different tissue types - across all species this is bigger pattern than shifts across age  
```{r, echo=FALSE, message=FALSE}
plot_Acc_by_Spp <- function(spp) {
data2 <- filter(SummarySppAge, species==spp)
plot(PerPropagule_mean ~ age, data=data2, xlab="age", ylab="proportion tissue", cex.axis=.8, las=1,
       main=labels.spp[spp], ylim=c(0,1), xlim=c(1,35), log="x", pch=16,col="orange")
points(PerPrePol_A_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8, col="red")
points(PerPrePol_S_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=17, col="blue")
points(PerPostPol_A_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=18, col="black")
points(PerPackDisp_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=19, col="light green")
legend(.9,.7, pch=c(16,8,17,18,19),legend=c("Propagule","Aborted pre-poll","Successful pre-poll","Aborted post-poll","Packaging & disp"), col=c("orange","red","blue","black","light green"), cex=.8, bty="n")
}

@@

###Considering relative effect of different accessory costs on RA
- Using models that look at all the various factors at once - but need to include species in these models and probably use some sort of a MANOVA  
- also issues because significance of various variables changes enormously with transformations  
```{r, echo=FALSE, message=FALSE}
mod <- lm(asin(sqrt(RA))~asin(sqrt(PerPropagule))+asin(sqrt(PerPrePol_A))+ asin(sqrt(PerPrePol_S))+ asin(sqrt(PerPostPol_A))+ asin(sqrt(PerPackDisp)) + species,data=subset(SummaryInd,Total>0))

mod <- lm(asin(sqrt(RA))~asin(sqrt(PerPropagule))+asin(sqrt(PerPrePol_A))+ asin(sqrt(PerPrePol_S))+ asin(sqrt(PerPostPol_A))+ asin(sqrt(PerPackDisp)) +asin(sqrt(PerPropagule))*species+asin(sqrt(PerPrePol_A))*species+ asin(sqrt(PerPrePol_S))*species+ asin(sqrt(PerPostPol_A))*species+ asin(sqrt(PerPackDisp))*species+ species,data=subset(SummaryInd,Total>0))
summary(mod)
plot(mod)
step(mod)
```

>>>>

###How do accessory costs shift with seed size?
- Larger-seeded species have higher accessory costs. 
- Also shown in other studies and suggested as a mechanism for why angiosperms were able to have such small seeds (relative to gymnosperms, with no decrease in proportional accessory costs - with decreasing size)  
```{r, echo=FALSE,message=FALSE}
yvar <- "prop_accessory_mean"
xvar <- "seed_size"
data <- SummarySpp

plot_yvar_vs_xvar(data, yvar, xvar,log="x",
                  xlab=xvar, ylab=yvar)
mod <- lm(log(seed_size)~prop_accessory_mean,SummarySpp)
summary(mod)
anova(mod)
```

###Do species differ in the proportional energy allocated to propagules (vs. accessory tissues)  
- COER, EPMI, HEPU, LEES, PILI invest the most energy in propagules  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_propagule"
xvar <- "species"
data <- filter(SummaryInd,Total>0)

plot_yvar_vs_xvar(data, yvar, xvar,
                 xlab=xvar, ylab=yvar)
mod <- lm(asin(sqrt(prop_propagule)) ~ species-1, data)
summary(mod)
data.frame(
    select(glance(mod), r.squared, p.value),
    n= length(resid(mod)),
    a = coef(mod)[1],
    b = coef(mod)[2]
  )

tmp <- summary(mod)
tmp <- tmp$coefficients
tmp <-as.data.frame(tmp)
tmp <- cbind(names(labels.spp),tmp)
names(tmp) <- c("species","estimate.trans","se.trans","t.value","p.value")
tmp2 <- select(SummarySpp, species, prop_propagule_mean, prop_propagule_se)
tmp <- merge(tmp, tmp2, by="species")
tmp
plot(prop_propagule_mean~species,tmp,cex.axis=.8,pch=16,log="y",ylim=c(.001,.25),main="Percent reproductive investment put into propagules")

mean1 <- tmp$prop_propagule_mean
se1 <- tmp$prop_propagule_se

@@

#for(i in 1:14) {
#arrows(i,tmp$prop_propagule_mean[i]+tmp$prop_propagule_se[i],i,tmp$prop_propagule_mean[i]-tmp$prop_propagule_se[i],code=3,angle=90, length=0.1)}

#barplot(height=test$prop_propagule_mean, names=test$species,cex.axis=.8,ylim=c(0,.25),cex.names=.8)
#xv <- barplot(height=test$prop_propagule_mean, names=test$species,cex.axis=.8,ylim=c(0,.25),cex.names=.8)
#xv

#points((prop_propagule_mean+prop_propagule_se)~species,test,pch=25)
#points((prop_propagule_mean-prop_propagule_se)~species,test,pch=24)

@@

###Do species differ in the proportional energy allocated to aborted pre-pollination tissues
- BOLE, EPMI, GRSP, HATE invest the most energy in aborted pre-pollination tissues  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_prepollen_aborted"
xvar <- "species"
data <- filter(SummaryInd,Total>0 & age>2)

plot_yvar_vs_xvar(data, yvar, xvar,
                  xlab=xvar, ylab=yvar)
mod <- lm(asin(sqrt(prop_prepollen_aborted)) ~ species, data)
summary(mod)
```

##Do individual species decrease accessory costs as they increase in size, age or RA?
###Change in accesory costs With size:  
- prop_accessory is arcsine transformed, size is log transformed  
- COER, HATE, and PILI show a significant decrease in accessory costs with increasing size (p<0.05)  
- GRSP shows a marginally significant decrease in accessory costs with increasing size (0.05<p<0.10)  
- PEPU shows a marginally significant INCREASE in accessory costs with increasing size (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_accessory"
xvar <- "total_weight"

@@

###Change in accesory costs With age:
- prop_accessory is arcsine transformed, age is log transformed  
- BOLE, COER, GRSP, HATE, HEPU, PELA show a significant decrease in accessory costs with increasing age (p<0.05)  
- EPMI, GRBU shows a marginally significant decrease in accessory costs with increasing age (0.05<p<0.10)  
- PUTU shows a sig INCREASE in accessory costs with increasing age  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_accessory"
xvar <- "age"

@@

###Change in accesory costs With respect to RA:  
- both RA and prop_accessory are arcsine transformed  
- BAER, COER, GRBU, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in accessory costs with increasing RA (p<0.05)  
- BOLE shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_propagule"
xvar <- "RA"

@@

##Next considering if the amount of energy spent on aborted buds shifts with age, size, RA  
###Change in aborted bud investment with size:  
- prop_prepollen_aborted is arcsine transformed, size is log transformed  
- GRSP, HATE, HEPU show a significant decrease in aborted buds with increasing size (p<0.05)  
- PUTU shows a marginally significant decrease in accessory costs with increasing RA (p ~ 0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_prepollen_aborted"
xvar <- "total_weight"

@@

###Change in aborted bud investment with respect to age:  
- prop_prepollen_aborted is arcsine transformed, age is log transformed  
- BAER, BOLE, COER, EPMI, LEES, PHPH, PILI show a significant decrease in aborted buds with increasing age (p<0.05)  
- GRSP, HATE show a marginally significant decrease in aborted buds with increasing age (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_prepollen_aborted"
xvar <- "age"

@@

###Change in aborted bud investment with respect to RA:
- both RA and prop_prepollen_aborted are arcsine transformed  
- BAER, BOLE, COER, GRSP, HATE, LEES, PELA, PEPU show a significant decrease in aborted buds with increasing RA (p<0.05)  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_prepollen_aborted"
xvar <- "RA"

@@

##Do individual species decrease total pre-pollination costs as they increase in size, age or RA?  
###Do total pre-pollination costs change with respect to size:  
- prop_prepollen_all is arcsine transformed, size is log transformed  
- GRSP, HATE show a significant decrease in total pre-pollination costs with increasing size (p<0.05)  
- HEPU shows a marginally significant decrease in total pre-pollination costs with increasing size (0.05<p<0.10)

```{r, echo=FALSE, message=FALSE}
yvar <- "prop_prepollen_all"
xvar <- "total_weight"

data <- filter(SummaryInd,ReproInv>0)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

@@

###Do total pre-pollination costs change with respect to age:
- prop_prepollen_all is arcsine transformed, age is log transformed  
- BAER, BOLE, COER, HEPU, LEES, PILI show a significant decrease in total pre-pollination costs with increasing age (p<0.05)  
- EPMI, GRBU, GRSP, HATE shows a marginally significant decrease in total pre-pollination costs with increasing age (0.05<p<0.10)  
- PUTU shows a sig INCREASE in total pre-pollination  costs with increasing age  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_prepollen_all"
xvar <- "age"

data <- filter(SummaryInd,ReproInv>0&age>2)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

@@

### Do total pre-pollination costs change with respect to RA:  
- both RA and prop_prepollen_all are arcsine transformed    
- BAER, BOLE, COER, GRSP, HATE, HEPU, LEES, PEPU show a significant decrease in total pre-pollination costs with increasing RA (p<0.05)  
- PELA shows a marginally significant decrease in total pre-pollination costs with increasing RA (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_prepollen_all"
xvar <- "RA"

data <- filter(SummaryInd,ReproInv>0&age>2)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  asin(sqrt(x[[xvar]]))))
results <- summarise_all_fits(lmfits)

@@

### Do total pre-pollination costs change with respect to Reproductive Investment:
- prop_prepollen_all is arcsine transformed, ReproInv is log transformed  
- COER, GRBU, GRSP, HATE, PEPU, PHPH, PILI show a significant decrease in total pre-pollination costs with increasing reproductive investment (p<0.05)  
- LEES shows a marginally significant decrease in total pre-pollination costs with increasing reproductive investment (0.05<p<0.10)  
```{r, echo=FALSE, message=FALSE}
yvar <- "prop_prepollen_all"
xvar <- "ReproInv"

data <- filter(SummaryInd,ReproInv>0&age>2)
data <- split(data, data$species)
lmfits <- lapply(data, function(x) lm(asin(sqrt(x[[yvar]])) ~  log(x[[xvar]])))
results <- summarise_all_fits(lmfits)

@@

###For each species, look at pattern of investment in different tissues across ages  
- Most obvious pattern is that different species invest more or less in different tissue types - across all species this is bigger pattern than shifts across age  
```{r, echo=FALSE, message=FALSE}
plot_Acc_by_Spp <- function(spp) {
data2 <- filter(SummarySppAge, species==spp)
plot(prop_propagule_mean ~ age, data=data2, xlab="age", ylab="proportion tissue", cex.axis=.8, las=1,
       main=labels.spp[spp], ylim=c(0,1), xlim=c(1,35), log="x", pch=16,col="orange")
points(prop_prepollen_aborted_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=8, col="red")
points(prop_prepollen_success_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=17, col="blue")
points(prop_postpollen_aborted_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=18, col="black")
points(prop_packaging_dispersal_mean ~ age, data=subset(SummarySppAge, species==spp), cex=1.3, pch=19, col="light green")
legend(.9,.7, pch=c(16,8,17,18,19),legend=c("Propagule","Aborted pre-poll","Successful pre-poll","Aborted post-poll","Packaging & disp"), col=c("orange","red","blue","black","light green"), cex=.8, bty="n")
}

@@

###Considering relative effect of different accessory costs on RA
- Using models that look at all the various factors at once - but need to include species in these models and probably use some sort of a MANOVA  
- also issues because significance of various variables changes enormously with transformations  
```{r, echo=FALSE, message=FALSE}
mod <- lm(asin(sqrt(RA))~asin(sqrt(prop_propagule))+asin(sqrt(prop_prepollen_aborted))+ asin(sqrt(prop_prepollen_success))+ asin(sqrt(prop_postpollen_aborted))+ asin(sqrt(prop_packaging_dispersal)) + species,data=subset(SummaryInd,Total>0))

mod <- lm(asin(sqrt(RA))~asin(sqrt(prop_propagule))+asin(sqrt(prop_prepollen_aborted))+ asin(sqrt(prop_prepollen_success))+ asin(sqrt(prop_postpollen_aborted))+ asin(sqrt(prop_packaging_dispersal)) +asin(sqrt(prop_propagule))*species+asin(sqrt(prop_prepollen_aborted))*species+ asin(sqrt(prop_prepollen_success))*species+ asin(sqrt(prop_postpollen_aborted))*species+ asin(sqrt(prop_packaging_dispersal))*species+ species,data=subset(SummaryInd,Total>0))
summary(mod)
plot(mod)
step(mod)
```
